---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

library(lwgeom)
library(rgdal)
library(reproj)

```
# PROJ

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![Travis build status](https://travis-ci.org/mdsumner/PROJ.svg?branch=master)](https://travis-ci.org/mdsumner/PROJ)
[![OSX Build Status](http://badges.herokuapp.com/travis/mdsumner/PROJ?branch=master&env=BUILD_NAME=osx_release&label=osx)](https://travis-ci.org/mdsumner/PROJ)
[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/mdsumner/PROJ?branch=master&svg=true)](https://ci.appveyor.com/project/mdsumner/PROJ)
[![Codecov test coverage](https://codecov.io/gh/mdsumner/PROJ/branch/master/graph/badge.svg)](https://codecov.io/gh/mdsumner/PROJ?branch=master)
[![CRAN status](https://www.r-pkg.org/badges/version/PROJ)](https://cran.r-project.org/package=PROJ)
[![CRAN_Download_Badge](http://cranlogs.r-pkg.org/badges/PROJ)](https://cran.r-project.org/package=PROJ
<!-- badges: end -->

The goal of PROJ is to provide generic coordinate system transformations and overcome some current
challenges and limitations in R. 


## Installation

WIP

# Notes


THINGS TO WORRY ABOUT:

* the *name* of this package
* t and z
* threading, see the PJ_CONTEXT
* coordinate order

https://proj4.org/development/quickstart.html

## Example

Minimal code example, two lon-lat coordinates to LAEA, and back. 


```{r minimal}
library(PROJ)
lon <- c(0, 147)
lat <- c(0, -42)
dst <- "+proj=laea +datum=WGS84 +lon_0=147 +lat_0=-42"

## forward transformation
(xy <- proj_trans(dst, lon, lat, INV = FALSE))

## inverse transformation
proj_trans(dst, xy[,1L], xy[,2L], xy[,3], INV = TRUE)


```

A more realistic example with coastline map data. 


```{r example}
library(PROJ)
w <- PROJ:::worlddata
lon <- na.omit(w[,1])
lat <- na.omit(w[,2])
dst <- "+proj=laea +datum=WGS84 +lon_0=147 +lat_0=-42"
xyz <- proj_trans(dst, X = lon, Y = lat, Z = rep(0, length(lon)), INV = FALSE)
plot(xyz, pch = ".")

lonlat <- proj_trans(dst, X = xyz[,1], Y = xyz[,2], INV = TRUE)
plot(lonlat, pch = ".")
```



# Speed comparisons

```{r speed}
library(reproj)
library(rgdal)
library(lwgeom)
library(sf)
lon <- w[,1]
lat <- w[,2]
lon <- rep(lon, 5)
lat <- rep(lat, 5)
ll <- cbind(lon, lat)
z <- rep(0, length(lon))
llproj <- "+proj=longlat +datum=WGS84"
# stll <- sf::st_crs(llproj)
# sfx <- sf::st_sfc(sf::st_multipoint(ll), crs = stll)  
rbenchmark::benchmark(
          PROJ = proj_trans(dst, lon, lat, z, INV = FALSE, quiet = TRUE), 
          reproj = reproj(cbind(lon, lat, z), target = dst, source = llproj), 
          rgdal = project(ll, dst), 
          sf_project = sf_project(llproj, dst, ll),
        # lwgeom = st_transform_proj(sfx, dst), 
        # sf = st_transform(sfx, dst), 
        replications = 100) %>% 
  dplyr::arrange(elapsed) %>% dplyr::select(test, elapsed, replications)
```

The comparison with rgdal is not exactly stunning, but with PROJ we can also do 3D transformations and (eventually, if needed) datum transformations and use the pipeline and other new features of PROJ (PROJ.4). 

```{r geocentric}
xyz <- proj_trans("+proj=geocent +datum=WGS84", lon, lat, z, INV = FALSE)
plot(as.data.frame(xyz), pch = ".")
```

Geocentric transformations aren't used in R much, but some examples are found in the [quadmesh](https://CRAN.R-project.org/package=quadmesh) and [anglr](https://github.com/hypertidy/anglr) packages. 

## Why PROJ?


The [reproj](https://CRAN.R-project.org/package=reproj) package wraps the very efficient `proj4::ptransform()` function for general coordinate system transformations.  Several package now use reproj for its consistency (no format or plumbing issues) and efficiency (directly transforming bulk coordinates). The proj4 package used by reproj doesn't provide the modern features of PROJ (PROJ.4), has not been updated on CRAN since 2012 and has an uncertain future. So reproj requires a new wrapper around PROJ (PROJ.4) itself. 

Since the 1990s [PROJ.4](https://proj4.org) has been the name of the common standard library for general coordinate system transformations (for geospatial). That 1994 release has been modernized and now has versions **PROJ 5** and **PROJ 6**. There's a bit of traction in the name PROJ.4, so it has stuck 

I'll use "PROJ (PROJ.4)" to distinguish the [system library](https://proj4.org) from [this package](https://github.com/mdsumner/PROJ).

There are a few links to the PROJ (PROJ.4) library in R. 

* [rgdal](https://CRAN.R-project.org/package=rgdal) Provides low level `project(matrix, inv = TRUE/FALSE)`, and engine behind `sp::spTransform()`. 
* [proj4](https://CRAN.R-project.org/package=proj4) Provides low level `project()` and `ptransform`. 
* [sf](https://CRAN.R-project.org/package=sf) Provides low level `sf::sf_project()` transformation of matrices. Provides high level `st_transform()` which works via the GDAL library and its own internal version of PROJ (PROJ.4). The high level function converts coordinates in list heirarchies of matrices into WKB for the transformations. 
* [lwgeom](https://CRAN.R-project.org/package=lwgeom) Provides high level `st_transform_proj()` also converts coordinates in list heirarchies of matrices into WKB, but internally uses the PROJ (PROJ.4) library directly. 

(The [mapproject](https://CRAN.R-project.org/package=mapproject) package uses all its own internal code). 

Packages sf, rgdal and proj4 provide raw access to coordinate transformations for R vectors. `sf::sf_project()` is the winner, but is embedded in a package that does many other things. Rgdal only has project forward and project inverse and always works in degrees proj4 has the more general `ptransform()` but requires manual conversion of degree values into radians. PROJ (PROJ.4) internally works only with radians. 

The rgdal function `project()` won't transform with a third Z coordinate. The sf functions do work with geocentric coords. 

The packages rgdal, sf, lwgeom are now compatible with PROJ 5 (and 6) and don't need any further attention in this regard. They work fine within their chosen context. 

Are there any other wrappers around PROJ (PROJ.4) on CRAN or Bioconductor, or in the works? Let me know! 


---

Please note that the 'PROJ' project is released with a
  [Contributor Code of Conduct](CODE_OF_CONDUCT.md).
  By contributing to this project, you agree to abide by its terms.
