---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

library(lwgeom)
library(rgdal)
library(reproj)

```
# PROJ


<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![R build status](https://github.com/hypertidy/PROJ/workflows/R-CMD-check/badge.svg)](https://github.com/hypertidy/PROJ/actions)
[![R build status](https://github.com/hypertidy/PROJ/workflows/test-coverage/badge.svg)](https://github.com/hypertidy/PROJ/actions)
[![R build status](https://github.com/hypertidy/PROJ/workflows/pkgdown/badge.svg)](https://github.com/hypertidy/PROJ/actions)
[![CRAN status](https://www.r-pkg.org/badges/version/PROJ)](https://cran.r-project.org/package=PROJ)
[![CRAN_Download_Badge](http://cranlogs.r-pkg.org/badges/PROJ)](https://cran.r-project.org/package=PROJ)


[![Travis NOPROJ](https://img.shields.io/travis/hypertidy/PROJ.svg?branch=master&env=BUILD_NAME=proj0&label=PROJ0)](https://travis-ci.org/hypertidy/PROJ)  no PROJ available 👍
<br>
[![Travis PROJ4](https://img.shields.io/travis/hypertidy/PROJ.svg?branch=master&env=BUILD_NAME=proj4&label=PROJ4)](https://travis-ci.org/hypertidy/PROJ) PROJ.4 in system, no function 👍
<br>
[![Travis PROJ5](https://img.shields.io/travis/hypertidy/PROJ.svg?branch=master&env=BUILD_NAME=proj5&label=PROJ5)](https://travis-ci.org/hypertidy/PROJ) PROJ 5 in system, no function 👍
<br>
[![Travis PROJ6](https://img.shields.io/travis/hypertidy/PROJ.svg?branch=master&env=BUILD_NAME=proj6&label=PROJ6)](https://travis-ci.org/hypertidy/PROJ) PROJ version 6, full function 🚀
<br>
[![Travis PROJ7](https://img.shields.io/travis/hypertidy/PROJ.svg?branch=master&env=BUILD_NAME=proj7&label=PROJ7)](https://travis-ci.org/hypertidy/PROJ) PROJ version 7, full function 🤸
<!-- badges: end -->

The goal of PROJ is to provide generic coordinate system transformations in R with a functional requirement for the system library PROJ >= 6. 

This is a shared goal with the [reproj](https://cran.r-project.org/package=reproj)
package, and PROJ provides the infrastructure for later versions of the underlying library. Reproj
otherwise uses the [proj4](https://cran.r-project.org/package=roj4) package, for library versions less than 6. 

PROJ provides basic coordinate transformations for generic numeric data in matrices or data frames. Transforming spatial data coordinates is a basic task, and has nothing to do with storage format.

PROJ is strictly for version 6 or higher of the PROJ library. The intention
is that this package will be used for when that version is available, and this
package can be compiled and installed even when it cannot do anything. For older
versions of PROJ (5, and 4) we can use the proj4 package.

Because we are version 6 or above only, there is no forward/inverse
transformation, only integrated source/target idioms. This is the same approach
taken by the reproj package- the source must be provided as well as the target.
When a data set has an in-built CRS projection recorded, then methods can be
written for that use-case with that format.

We can use "auth:code" forms, PROJ.4 strings, full WKT2, or the name of a CRS as
found in the PROJ database, e.g “WGS84”, “NAD27”, etc. Full details are provided
in the [PROJ
documentation](https://proj.org/development/reference/functions.html#c.proj_create).


## Things to be aware of

Note that for PROJ5 (and lower) this package is non-functional (can use proj4).
The testing here on Travis ensures that the package installs successfully for
various versions of PROJ, although underlying functionality is disabled for 4
and 5 (and for no PROJ).

* Input can be a data frame or a matrix, but internally input is assumed to be x, y, z, *and time*. So the output is always a 4-column matrix. 
* You can't use strings like "+init=epsg:4326" any more, it must be "epsg:4326". 
* You should know what your target projection is, and also what your source projection is. This is your responsibility. 

Personally, I need this low-level package in order to develop other projects. I
don't care about the big snafu regarding changes in version 6 and whatever, we
should have low-level tools and then we can code at the R level to sort stuff
out. A text-handler for various versions and validations of CRS representations
would be good, for instance we can just gsub out "+init=" for those sorts of
things, and being able to write "WGS84" as a valid source or target is a massive
bonus.

## WHY

A quick list, see more below. 

* Why not proj4?  It's not maintained in a way that works for me. 
* Why not sf?  It brings a lot of baggage, and can't do geocentric transformations. 
* Why not rgdal? Still baggage, no transformations possible without special data formats, no geocentric. 
* Why not lwgeom? That package is format-specific, and does not work with generic
data coordinates so is unsuitable for many straightforward and efficient
data-handling schemes.
* Why not mapproj? This is unusable for real-world projections in my experience,
it seems to be written for some basic graphics cases.
* Why not reproj? reproj will be improved by importing PROJ. This is an
extension for reproj, to bridge it from PROJ version 4 and 5, to version 6 and 7
and beyond.


## Installation

On Windows and MacOS, just do 

```R
install.packages("PROJ")
```

or 

```R
remotes::install_cran("PROJ")
```

to install the *static* binary version of the package. On Windows this ends in .zip, on MacOS it's a .tgz. *Static* means
all the stuff is in the package, there's no pre-requirement installations. 

For a brew installation on MacOS, first do 

```
brew install proj
```

WIP - see the matrix set up in .travis.yml, and the scripts in ci/travis/ - much
gratitude to GDAL for examples of how to do all this!  

See ./.github/workflows/check-standard.yaml for the Github Actions approach, many thanks to James Balamuta. 


# Notes

None of these things are dealt with. 

* threading, see the PJ_CONTEXT
* coordinate order
* the zero value after transformation, it comes out like -3.19835e-15 (do we just zapsmall()?)

https://proj.org/development/quickstart.html

## Example

Minimal code example, two lon-lat coordinates to LAEA, and back. 

```{r minimal}
library(PROJ)
lon <- c(0, 147)
lat <- c(0, -42)
dst <- "+proj=laea +datum=WGS84 +lon_0=147 +lat_0=-42"
src <- "+proj=longlat +datum=WGS84"

## forward transformation
(xy <- proj_trans_generic( cbind(lon, lat), dst, source = src))

## inverse transformation
proj_trans_generic(cbind(xy$x_, xy$y_), src, source = dst)


## note that NAs propagate in the usual way
lon <- c(0, NA, 147)
lat <- c(NA, 0, -42)


```

A more realistic example with coastline map data. 


```{r example}
library(PROJ)
w <- PROJ::xymap
lon <- na.omit(w[,1])
lat <- na.omit(w[,2])
dst <- "+proj=laea +datum=WGS84 +lon_0=147 +lat_0=-42"
xyzt <- proj_trans_generic(cbind(lon, lat), dst, source = "epsg:4326", z_ = 0)
plot(xyzt$x_, xyzt$y_, pch = ".")

lonlat <- proj_trans_generic(xyzt, src, source = dst)
plot(lonlat$x_, lonlat$y_, pch = ".")
```

## Convert projection strings

We can generate PROJ or within limitations WKT2 strings. 

```{r WKT2}
cat(wkt2 <- proj_create("EPSG:4326"))

## this is not proper WKT2 and cannot be used, compare to sf::st_crs("+proj=etmerc +lat_0=38 +lon_0=125 +ellps=bessel")
cat(proj_create("+proj=etmerc +lat_0=38 +lon_0=125 +ellps=bessel"))

proj_create(wkt2, format = 1L)


```


# Speed comparisons

```{r speed, warning=FALSE}
library(reproj)
library(rgdal)
library(lwgeom)
library(sf)
lon <- w[,1]
lat <- w[,2]
lon <- rep(lon, 25)
lat <- rep(lat, 25)
ll <- cbind(lon, lat)
z <- rep(0, length(lon))
llproj <- "+proj=longlat +datum=WGS84"

xyz <- cbind(lon, lat, z)
xyzt <- cbind(lon, lat, z, 0)
# stll <- sf::st_crs(llproj)
# sfx <- sf::st_sfc(sf::st_multipoint(ll), crs = stll) 

rbenchmark::benchmark(
          PROJ = proj_trans_generic(ll, target = dst, source = llproj, z_ = z),
          reproj = reproj(ll, target = dst, source = llproj),
          rgdal = project(ll, dst),
          sf_project = sf_project(llproj, dst, ll),
        # lwgeom = st_transform_proj(sfx, dst),
        # sf = st_transform(sfx, dst),
        replications = 100) %>%
  dplyr::arrange(elapsed) %>% dplyr::select(test, elapsed, replications)
```

The speed is not exactly stunning, but with PROJ we can also do 3D transformations and that's good enough for me.  I think it will be faster with the underlying API function `proj_trans_array()`, instead of `proj_trans_generic()`, but I don't really know. 

A geocentric example, suitable for plotting in rgl. 

```{r geocentric}
xyzt <- proj_trans_generic(cbind(w[,1], w[,2]), target = "+proj=geocent +datum=WGS84", source = "EPSG:4326")
plot(as.data.frame(xyzt[1:3]), pch = ".", asp = 1)
```

Geocentric transformations aren't used in R much, but some examples are found in the [quadmesh](https://CRAN.R-project.org/package=quadmesh) and [anglr](https://github.com/hypertidy/anglr) packages. 

## Why PROJ?


The [reproj](https://CRAN.R-project.org/package=reproj) package wraps the very efficient `proj4::ptransform()` function for general coordinate system transformations.  Several package now use reproj for its consistency (no format or plumbing issues) and efficiency (directly transforming bulk coordinates). The proj4 package used by reproj doesn't provide the modern features of PROJ (PROJ.4), has not been updated on CRAN since 2012 and has an uncertain future. So reproj requires a new wrapper around PROJ (PROJ.4) itself. 

Since the 1990s [PROJ.4](https://proj4.org) has been the name of the common standard library for general coordinate system transformations (for geospatial). That 1994 release has been modernized and now has versions **PROJ 5** and **PROJ 6**. There's a bit of traction in the name PROJ.4, so it has stuck 

I'll use "PROJ (PROJ.4)" to distinguish the [system library](https://proj4.org) from [this package](https://github.com/hypertidy/PROJ).

There are a few links to the PROJ (PROJ.4) library in R. 

* [rgdal](https://CRAN.R-project.org/package=rgdal) Provides low level `project(matrix, inv = TRUE/FALSE)`, and engine behind `sp::spTransform()`. 
* [proj4](https://CRAN.R-project.org/package=proj4) Provides low level `project()` and `ptransform`. 
* [sf](https://CRAN.R-project.org/package=sf) Provides low level `sf::sf_project()` transformation of matrices. Provides high level `st_transform()` which works via the GDAL library and its own internal version of PROJ (PROJ.4). The high level function converts coordinates in list heirarchies of matrices into WKB for the transformations. 
* [lwgeom](https://CRAN.R-project.org/package=lwgeom) Provides high level `st_transform_proj()` also converts coordinates in list hierarchies of matrices into WKB, but internally uses the PROJ (PROJ.4) library directly. 

(The [mapproj](https://CRAN.R-project.org/package=mapproj) package uses all its own internal code). 

Packages sf, rgdal and proj4 provide raw access to coordinate transformations for R vectors. `sf::sf_project()` is the winner, but is embedded in a package that does many other things. Rgdal only has project forward and project inverse and always works in degrees proj4 has the more general `ptransform()` but requires manual conversion of degree values into radians. PROJ (PROJ.4) internally works only with radians. 

The rgdal function `project()` won't transform with a third Z coordinate. The sf functions do work with geocentric coords. 

The packages rgdal, sf, lwgeom are now compatible with PROJ 5 (and 6) and don't need any further attention in this regard. They work fine within their chosen context. 

Are there any other wrappers around PROJ (PROJ.4) on CRAN or Bioconductor, or in the works? Let me know! 


---

Please note that the PROJ project is released with a
  [Contributor Code of Conduct](https://github.com/hypertidy/PROJ/blob/master/CODE_OF_CONDUCT.md).
  By contributing to this project, you agree to abide by its terms.
