---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

library(lwgeom)
library(rgdal)
library(reproj)

```
# PROJ


The goal of PROJ is to provide generic coordinate system transformations in R with a functional requirement for the system library to be provided by the [libproj package](https://cran.r-project.org/package=libproj). 

This is a shared goal with the [reproj](https://cran.r-project.org/package=reproj)
package, and PROJ provides the infrastructure for later versions of the underlying library. 

PROJ provides basic coordinate transformations for generic numeric data in matrices or data frames. Transforming spatial data coordinates is a basic task independent of storage format.

PROJ is strictly for modern versions of the PROJ library, now via [libproj package](https://cran.r-project.org/package=libproj) at version 7. Alternative pathways to 
PROJ via the proj4 package are available but only used for expert testing.

Because we are using modern PROJ there is no forward/inverse
transformation, only integrated source/target idioms. This is the same approach
taken by the reproj package- the source must be provided as well as the target.
When a data set has an in-built CRS projection recorded, then methods can be
written for that use-case with that format.

We can use 'auth:code' forms, PROJ.4 strings, full WKT2, or the name of a CRS as
found in the PROJ database, e.g 'WGS84', 'NAD27', etc. Full details are provided
in the [PROJ
documentation](https://proj.org/development/reference/functions.html#c.proj_create).


## Things to be aware of

Note that *PROJ in your system* is not used, PROJ imports the standalone libproj package installed from CRAN. 

* Input can be a data frame or a matrix, but internally input is assumed to be x, y, z, *and time*. So the output is always a 4-column matrix. 
* You can't use strings like "+init=epsg:4326" any more, it must be "EPSG:4326". 
* You should know what your target projection is, and also what your source projection is. This is your responsibility. 
* PROJ assumes longitude/latitude order always by setting the PROJ library context *proj_normalize_for_visualization*. 

Please see [PROJ library documentation](https://proj.org/development/quickstart.html) for details on this. 



## Installation

On all systems do 

```R
install.packages("PROJ")
```

or 

```R
remotes::install_cran("PROJ")
```

To install the development version from Github do

```R
remotes::install_github("hypertidy/PROJ")
```


## Example

Minimal code example, two lon-lat coordinates to LAEA, and back. 

```{r minimal}
library(PROJ)
lon <- c(0, 147)
lat <- c(0, -42)
dst <- "+proj=laea +datum=WGS84 +lon_0=147 +lat_0=-42"
src <- "+proj=longlat +datum=WGS84"

## forward transformation
(xy <- proj_trans( cbind(lon, lat), dst, source = src))

## inverse transformation
proj_trans(cbind(xy$x_, xy$y_), src, source = dst)


## note that NAs propagate in the usual way
lon <- c(0, NA, 147)
lat <- c(NA, 0, -42)

proj_trans(cbind(lon, lat), src, source = dst)

```

A more realistic example with coastline map data. 


```{r example}
library(PROJ)
w <- PROJ::xymap
lon <- na.omit(w[,1])
lat <- na.omit(w[,2])
dst <- "+proj=laea +datum=WGS84 +lon_0=147 +lat_0=-42"
xy <- proj_trans(cbind(lon, lat), dst, source = "epsg:4326")
plot(xy$x_, xy$y_, pch = ".")

lonlat <- proj_trans(xy, src, source = dst)
plot(lonlat$x_, lonlat$y_, pch = ".")
```

## Convert projection strings

We can generate PROJ or within limitations WKT2 strings, format 0, 1, 2 for WKT, proj4string, projjson respectively. 

```{r WKT2, eval=TRUE}
cat(wkt2 <- proj_crs_text("EPSG:4326"))

proj_crs_text(wkt2, format = 1L)


```


# Speed comparisons

```{r speed, warning=FALSE}
library(reproj)
library(rgdal)
library(lwgeom)
library(sf)
lon <- w[,1]
lat <- w[,2]
lon <- rep(lon, 25)
lat <- rep(lat, 25)
ll <- cbind(lon, lat)
z <- rep(0, length(lon))
llproj <- "+proj=longlat +datum=WGS84"

xyz <- cbind(lon, lat, z)
xyzt <- cbind(lon, lat, z, 0)

rbenchmark::benchmark(
          PROJ = proj_trans(ll, target = dst, source = llproj),
          reproj = reproj::reproj(ll, target = dst, source = llproj),
          rgdal = project(ll, dst),
          sf_project = sf_project(llproj, dst, ll),
        # lwgeom = st_transform_proj(sfx, dst),
        # sf = st_transform(sfx, dst),
        replications = 100) %>%
  dplyr::arrange(elapsed) %>% dplyr::select(test, elapsed, replications)
```



A geocentric example, suitable for plotting in rgl. 

```{r geocentric}
xyzt <- proj_trans(cbind(w[,1], w[,2]), z_ = rep(0, dim(w)[1L]), target = "+proj=geocent +datum=WGS84", source = "EPSG:4326")
plot(as.data.frame(xyzt[1:3]), pch = ".", asp = 1)
```

Geocentric transformations aren't used in R much, but some examples are found in the [quadmesh](https://CRAN.R-project.org/package=quadmesh) and [anglr](https://github.com/hypertidy/anglr) packages. 

## Why PROJ?

PROJ was created before [libproj](https://cran.r-project.org/package=libproj) existed, and now leverages it. 

* Why not proj4?  It's not maintained in a way that works for me. 
* Why not sf?  It brings a lot of baggage, and can't do geocentric transformations. 
* Why not rgdal? Still baggage, no transformations possible without special data formats, no geocentric. 
* Why not lwgeom? That package is format-specific, and does not work with generic
data coordinates so is unsuitable for many straightforward and efficient
data-handling schemes.
* Why not mapproj? This is unusable for real-world projections in my experience,
it seems to be written for some basic graphics cases.
* Why not reproj? reproj will be improved by importing PROJ. This is an
extension for reproj, to bridge it from PROJ version 4 and 5, to version 6 and 7
and beyond.

The [reproj](https://CRAN.R-project.org/package=reproj) package wraps the very efficient `proj4::ptransform()` function for general coordinate system transformations.  Several package now use reproj for its consistency (no format or plumbing issues) and efficiency (directly transforming bulk coordinates). The proj4 package is not updated regularly, and depends on a local installation of the system library on some sytems. . So reproj requires a new wrapper around PROJ (PROJ.4) itself. 

Since the 1990s [PROJ.4](https://proj4.org) has been the name of the common standard library for general coordinate system transformations (for geospatial). That 1994 release has been modernized and now has versions **PROJ 5** and **PROJ 6**. There's a bit of traction in the name PROJ.4, so it has stuck 

I'll use "PROJ (PROJ.4)" to distinguish the [system library](https://proj4.org) from [this package](https://github.com/hypertidy/PROJ).

There are a few links to the PROJ (PROJ.4) library in R. 

* [rgdal](https://CRAN.R-project.org/package=rgdal) Provides low level `project(matrix, inv = TRUE/FALSE)`, and engine behind `sp::spTransform()`. 
* [proj4](https://CRAN.R-project.org/package=proj4) Provides low level `project()` and `ptransform`. 
* [sf](https://CRAN.R-project.org/package=sf) Provides low level `sf::sf_project()` transformation of matrices. Provides high level `st_transform()` which works via the GDAL library and its own internal version of PROJ (PROJ.4). The high level function converts coordinates in list heirarchies of matrices into WKB for the transformations. 
* [lwgeom](https://CRAN.R-project.org/package=lwgeom) Provides high level `st_transform_proj()` also converts coordinates in list hierarchies of matrices into WKB, but internally uses the PROJ (PROJ.4) library directly. 

(The [mapproj](https://CRAN.R-project.org/package=mapproj) package uses all its own internal code). 

Packages sf, rgdal and proj4 provide raw access to coordinate transformations for R vectors. `sf::sf_project()` is the winner, but is embedded in a package that does many other things. Rgdal only has project forward and project inverse and always works in degrees proj4 has the more general `ptransform()` but requires manual conversion of degree values into radians. PROJ (PROJ.4) internally works only with radians. 

The rgdal function `project()` won't transform with a third Z coordinate. The sf functions do work with geocentric coords. 

The packages rgdal, sf, lwgeom are now compatible with PROJ 5 (and 6) and don't need any further attention in this regard. They work fine within their chosen context. 


---

Please note that the PROJ project is released with a
  [Contributor Code of Conduct](https://github.com/hypertidy/PROJ/blob/master/CODE_OF_CONDUCT.md).
  By contributing to this project, you agree to abide by its terms.
